---
title: "reflection of Laplace for phd in USA"
format: pdf
editor:
  Tianrui_Fu: {}
  Yiyue_Deng: {}
  Dawson_Li: {}
---

```{r}
#| warning: false
#| message: false
#| echo: false
library(haven)
library(tidyverse)
library(labelled)
ipums_extract <- read_csv("/Users/frankstrove/Downloads/Archive/usa_00001.csv")


ipums_extract <- 
  ipums_extract |>
  select(STATEICP, EDUCD) |>
  to_factor()
```

Making use of the codebook, how many respondents were there in each state (STATEICP) that had a doctoral degree as their highest educational attainment (EDUC)? 

```{r}
#| warning: false
#| message: false
#| echo: false
library(knitr)
library(dplyr)
library(kableExtra)

doctoral_degree_count <- ipums_extract |>
  filter(EDUCD == 116) |>  # Filter rows with "Doctoral degree"
  group_by(STATEICP) |>
  summarise(doctoral_count = n()) |>  # Count respondents with a doctoral degree in each state
  ungroup()

kable(doctoral_degree_count, format = "markdown", caption = "Count of Doctoral Degree Holders by State") |> 
  kable_styling(full_width = F, position = "left")
```

# Instructions for obtaining the data.

To obtain data from IPUMS USA, we first navigated to the IPUMS website and selected "IPUMS USA." Next, we clicked on "Get Data" and chose "2022 ACS" under the "SELECT SAMPLE" section. For state-level data, we selected "HOUSEHOLD," then "GEOGRAPHIC," and added "STATEICP" to our cart. For individual-level data, we selected "PERSON," then navigated to "DEMOGRAPHIC" and "EDUCATION" where we added "SEX" and "EDUC" to our cart respectively. Afterward, we clicked "VIEW CART" followed by "CREATE DATA EXTRACT." We changed the "DATA FORMAT" to ".csv" and clicked "SUBMIT EXTRACT." After creating an account and verifying it, received an email notification when the extract was ready. Finally, we downloaded and saved the file locally (e.g., "usa_00001.csv.gz") for use in `R`.

# A concise summary of the ratio estimators method.

Ratio estimation is a statistical method that estimates population totals or means by using known ratios from a sample. By calculating the ratio of a characteristic (e.g., individuals with a PhD) to the total population in a known subset (e.g., California), researchers can apply this ratio to other subsets to infer broader population totals. This approach is particularly useful when direct population measurement is impractical, enhancing estimation accuracy, especially when there’s a strong correlation between the variable of interest and an auxiliary variable.

# The actual number of respondents and your estimations.

```{r}
#| warning: false
#| message: false
#| echo: false
# Get the total count of respondents in California
total_respondents_california <- 391171  # Given value for California from Quercus-reflection 4 page

# Get the number of respondents with a doctoral degree in California
phd_respondents_cal <- doctoral_degree_count |>
  filter(STATEICP == "71") |>
  pull(doctoral_count)

# Calculate the Laplace ratio of doctoral degree holders to total respondents in California
phd_ratio_cal <- phd_respondents_cal / total_respondents_california

# Get the total count of respondents in each state using the ratio estimator
est_totalcounts <- doctoral_degree_count |>
  mutate(estimated_total = doctoral_count / phd_ratio_cal)

# Merge the estimated totals with actual respondent counts
actual_count <- ipums_extract |>
  group_by(STATEICP) |>
  summarise(actual_total = n()) |>  # Count actual number of respondents in each state
  ungroup()

compare_tbl <- doctoral_degree_count |>
  left_join(actual_count, by = "STATEICP") |>
  left_join(est_totalcounts, by = "STATEICP") |>
  select(STATEICP, actual_total, estimated_total)

# Display the compare_tbl between actual and estimated counts
kable(compare_tbl, format = "markdown", caption = "Compare of actual and estimated Holders by State") |> 
  kable_styling(full_width = F, position = "left")
```

```{r}
#| warning: false
#| message: false
#| echo: false
ggplot(compare_tbl, aes(x = STATEICP)) +
  geom_line(aes(y = actual_total, color = "Actual Total")) +  # Line for actual total
  geom_line(aes(y = estimated_total, color = "Estimated Total")) +  # Line for estimated total
  labs(title = "Comparison of Actual vs. Estimated Total Respondents by State",
       x = "State ICP",
       y = "Number of Respondents",
       color = "Legend") +
  theme_minimal()
```

# A brief reasoning for what you think that they differ.

The estimated total number of respondents in each state using the ratio estimators approach may differ from the actual number of respondents for several reasons:

-   Ratio estimation methods typically assume that relationships observed in one subset of the data are representative of other subsets. However, this assumption often does not hold in real world. States may have very different demographics, economic conditions, and educational policies, making it inaccurate to generalize patterns of educational attainment in one state to other states.

-   Educational attainment across U.S. states varies significantly due to differences in regional policies, cultural factors, and access to economic opportunities. Some states invest more in higher education, offering greater resources and opportunities, while others may prioritize vocational training or immediate entry into the workforce due to economic or cultural reasons. California, with its large population and abundant resources, represents a unique case where its educational patterns may not be generalizable to other states with differing economic conditions and cultural contexts. Thus, using California's ratio to estimate national trends could lead to inaccurate conclusions.

-   The ratio estimation method relies on the assumption that there is a consistent relationship between the characteristics of doctoral degrees and the total population in each state. If this relationship is inconsistent, the estimates derived from this method will be biased. For example, states with different levels of public education funding may have different relationships between population and educational attainment, resulting in biased estimates.

    These reasons indicate that the assumption of homogeneity used in ratio estimators often leads to differences when applied to diverse populations such as different states in the US.

# Reference
